### 2019-02-07 Sections/economics_of_the_quantum_internet.tex
SH corrections

To analyse the decision making options for qubit suppliers, game theory is an analytical tool to understand `games' between players, where the outcome of the game is dependent on the various strategies employed by the players. The most well-known of these games is the prisoner's dilemma\index{Prisoner's dilemma} \cite{(Flood and Dresher, 1950)}\comment{Please provide refs.}, describing the potential risks and rewards for two prisoners who are being independently questioned about a crime. This type of game is typically called a \textit{two-person, non-zero sum, non-cooperative game}\index{Two-person, non-zero sum, non-cooperative games} \cite{bib:Bacharach76}. More detailed examples of game theory have explored many of the base assumptions of this scenario, such as what if the prisoners are able to cooperate from the outset? How does this then result in maximising utility, and is cooperation always the best answer?

@@

For the uninitiated to economic analysis, particularly game theory, a few key concepts need to be established. This chapter by no means tries to cover these concepts in complete detail. For more detailed information we suggest referring to \cite{???}\comment{Please provide refs.}. Furthermore, the analysis at this point is only descriptive in nature, as a means of establishing the space where new research can be developed. Further more rigorous investigation is encouraged. The essential concepts that we rely on taken from game theory are:

@@

An individual user of a quantum computer operating on their own, observes computational power characterised by the appropriate computational scaling function acting on the number qubits in their possession, $f_\mathrm{sc}(n)$. This stipulates the utility payoff for that individual, allowing for trivial  construction (and efficient mathematical representation) of multi-player payoff matrices simply by characterising single-player payoffs independently.

@@

When implementing quantum computations characterised by completely general computational scaling functions\index{Computational scaling functions}, $f^{X,Y}_\mathrm{sc}$ (which in general can be distinct for the two players, $X$ and $Y$), when using an \textsc{Isolation} strategy, the respective payoff matrix elements are simply given by,
\begin{align}\label{eq:util_payoff_iso}
\textsc{Payoff} = (f^X_\mathrm{sc}(n_X),f^Y_\mathrm{sc}(n_Y)),
\end{align}
where payoff matrix elements are assumed to be in units of classical-equivalent processing time\footnote{There's nothing unique or special about using classical-equivalent computational power as our utility measure. Any other measure of `payoff' could be equally well justified, depending on circumstance. For example, one could instead represent utility in terms of the monetary value of computational power. In that case, we simply need to transform the payoff matrix elements using the cost of computation\index{Cost of computation} identity presented in Post.~\ref{post:cost_comp}.} (i.e FLOPs), and resource-sharing is based on the methodology for arbitrage-free time-share allocation presented in Sec.~\ref{sec:arb_free_time_share}.

@@

Elements associated with \textsc{License} strategies are transformed as,
\begin{align}
\begin{pmatrix}
   f^X_\mathrm{sc}(n_X) \\
   f^Y_\mathrm{sc}(n_Y)
\end{pmatrix} \underset{\textsc{License}}{\longrightarrow} \begin{pmatrix}
   (1-r) \cdot f^X_\mathrm{sc}(n_X) \\
   f^Y_\mathrm{sc}(n_Y) + r \cdot f^Y_\mathrm{sc}(n_X)
\end{pmatrix},
\end{align}
where \mbox{$0\leq r\leq 1$} denotes the proportion of $X$'s compute-time licensed to $Y$. We have here implicitly assumed that $Y$'s compute-time is fully exploited by $Y$ alone, and none is licensed in the reverse direction\footnote{It would be a fruitless exercise for both $X$ and $Y$ to simultaneously license the same commodity to one another in both directions. Indeed, in a more realistic setting, in which transaction and communication costs are accommodated for, this would be inefficient and positively counterproductive.}.

This is easily logically generalised to an arbitrary multi-player setting, in which case the transformation becomes,
\begin{align}
	f_\mathrm{sc}^i(n_i) \underset{\textsc{License}}{\longrightarrow} \sum_{j=1}^N r_{j\to i} f_\mathrm{sc}^i(n_j),
\end{align}
where there are $N$ players, all engaging with one another using licensing only. Here, $r_{i\to j}$ denotes the proportion of $i$'s compute-time licensed to $j$, which are normalised (assuming no wasted clock-cycles) for all users such that,
\begin{align}
\sum_{j=1}^N r_{i\to j} = 1\,\,\forall\,i.	
\end{align}
Note that when \mbox{$r_{i\to j}=\delta_{i,j}$} (i.e \mbox{$r=I_N$} is the \mbox{$N\times N$} identity matrix, and there is no inter-player licensing) the \textsc{License} strategy simply reduces back to the \textsc{Isolation} strategy.

@@

Elements associated with \textsc{Unify} strategies will undergo the quantum utility payoff enhancement,
\begin{align}\label{eq:util_payoff_enh_1}
(f^X_\mathrm{sc}(n_X),f^Y_\mathrm{sc}(n_Y)) \underset{\textsc{Unify}}{\longrightarrow} (c_X,c_Y),
\end{align}
where,
\begin{align}\label{eq:util_payoff_enh_2}
c_X &= \frac{n_X}{n_X+n_Y}\cdot f^X_\mathrm{sc}(n_X+n_Y),\nonumber\\
c_Y &= \frac{n_Y}{n_X+n_Y}\cdot f^Y_\mathrm{sc}(n_X+n_Y),
\end{align}
from the definition for time-shared compute power given in Def.~\ref{def:time_share_comp_power}.

As before, we can logically generalise the payoff enhancement of a generalised \textsc{Unify} strategy to the multi-player scenario as,
\begin{align}
	f_\mathrm{sc}^i(n_i) \underset{\textsc{Unify}}{\longrightarrow} \frac{n_i}{\sum_{j=1}^N n_j} \cdot f_\mathrm{sc}^i\left(\sum_{j=1}^N n_j\right).
\end{align}

Thus, it is evident that the enhancement in \textsc{Unify} strategies is highly dependent on:
\begin{itemize}
	\item The total number of qubits held between the players,
		\begin{align}
			\sum_j n_j.
		\end{align}
	\item The asymmetry in the distribution of qubits between them,
		\begin{align}
			\frac{n_i}{\sum_{j=1}^N n_j}.
		\end{align}
	\item The respective algorithms to which the computational resources are being applied by each player, which influence the subjective scaling functions, $f^i_\mathrm{sc}$, independently.
\end{itemize}

@@

What happens with collaboration between players who have different systems? If this assumption is relaxed, what comes of the resulting utility payoff matrix and choice of strategies? The analysis for this assumption had already been presented in the previous discussion. The effective impact of taxation is that it will limit the amount of computational power delivered to market and capture governmental revenue from the sale of compute-time. This is analogous to what happens if $X$ was to build a less powerful system than $Y$. Through collaboration, they could still attain point $Q'$, shown in Fig.~\ref{fig:game_theory_4}, but this time the asymmetry arises as a design consequence rather than a regulatory imposition. Despite $X$ now not contributing as much to the overall computational power of the quantum cloud, choosing the maximising strategy of offering all available computing power to the cloud does maximise individual utility. Furthermore, the subjective nature of the utility derivation means there doesn't have to be a change from the described payoff matrix shown in Fig.~\ref{fig:game_theory_3}. Despite having a less powerful system, the proportional allocation of financial rewards relative to the contribution, means both $X$ and $Y$ could maintain a symmetric utility payoff outcome, like in Fig.~\ref{fig:game_theory_3}. What can be concluded from this analysis is that in a 2-player cooperative game, the player with the more powerful system also has a greater incentive to cooperate\comment{Contradicts earlier conclusion. In my earlier derivations I argued that the leverage phenomena incentivises smaller players more than bigger ones since their leverage is greater. Please clarify or reconcile.}.

>>>>

To analyse the decision making options for qubit suppliers, game theory is an analytical tool to understand `games' between players, where the outcome of the game is dependent on the various strategies employed by the players. The most well-known of these games is the prisoner's dilemma\index{Prisoner's dilemma} \cite{bib:Poundstone93}, describing the potential risks and rewards for two prisoners who are being independently questioned about a crime.

Games can be classified based on their dimension, including the number of agents, the symmetry of the utility payoff, and whether they are cooperative. The Prisoner's dilemma is an example of a \textit{two-person, non-zero sum, non-cooperative game}\index{Two-person, non-zero sum, non-cooperative games} \cite{bib:Bacharach76}. More detailed examples of game theory have explored many of the base assumptions of this scenario, such as what if the prisoners are able to cooperate from the outset? How does this then result in maximising utility, and is cooperation always the best answer?

@@

For the uninitiated to economic analysis, particularly game theory, a few key concepts need to be established. This chapter by no means tries to cover these concepts in complete detail. For more detailed information we suggest referring to \cite{bib:Sugden04, bib:Bacharach76, bib:Straffin93}. Furthermore, the analysis at this point is only descriptive in nature, as a means of establishing the space where new research can be developed. Further more rigorous investigation is encouraged. The essential concepts that we rely on taken from game theory are:

@@

An individual user, $i$, of a quantum computer operating on their own, observes computational power characterised by the appropriate computational scaling function acting on the number qubits in their possession, $f_\mathrm{sc}^{(i)}(n)$. This stipulates the utility payoff for that individual, allowing for trivial  construction (and efficient mathematical representation) of multi-player payoff matrices simply by characterising single-player payoffs independently.

@@

When implementing quantum computations characterised by completely general computational scaling functions\index{Computational scaling functions}, $f^{(X,Y)}_\mathrm{sc}$ (which in general can be distinct for the two players, $X$ and $Y$), when using an \textsc{Isolation} strategy, the respective payoff matrix elements are simply given by,
\begin{align}\label{eq:util_payoff_iso}
\textsc{Payoff} = (f^{(X)}_\mathrm{sc}(n_X),f^{(Y)}_\mathrm{sc}(n_Y)),
\end{align}
where payoff matrix elements are assumed to be in units of classical-equivalent processing time\footnote{There's nothing unique or special about using classical-equivalent computational power as our utility measure. Any other measure of `payoff' could be equally well justified, depending on circumstance. For example, one could instead represent utility in terms of the monetary value of computational power. In that case, we simply need to transform the payoff matrix elements using the cost of computation\index{Cost of computation} identity presented in Post.~\ref{post:cost_comp}.} (i.e FLOPs), and resource-sharing is based on the methodology for arbitrage-free time-share allocation presented in Sec.~\ref{sec:arb_free_time_share}.

@@

Elements associated with \textsc{License} strategies are transformed as,
\begin{align}
\begin{pmatrix}
   f^{(X)}_\mathrm{sc}(n_X) \\
   f^{(Y)}_\mathrm{sc}(n_Y)
\end{pmatrix} \underset{\textsc{License}}{\longrightarrow} \begin{pmatrix}
   (1-r) \cdot f^{(X)}_\mathrm{sc}(n_X) \\
   f^{(Y)}_\mathrm{sc}(n_Y) + r \cdot f^{(Y)}_\mathrm{sc}(n_X)
\end{pmatrix},
\end{align}
where \mbox{$0\leq r\leq 1$} denotes the proportion of $X$'s compute-time licensed to $Y$. We have here implicitly assumed that $Y$'s compute-time is fully exploited by $Y$ alone, and none is licensed in the reverse direction\footnote{It would be a fruitless exercise for both $X$ and $Y$ to simultaneously license the same commodity to one another in both directions. Indeed, in a more realistic setting, in which transaction and communication costs are accommodated for, this would be inefficient and positively counterproductive.}.

This is easily logically generalised to an arbitrary multi-player setting, in which case the transformation becomes,
\begin{align}
	f_\mathrm{sc}^{(i)}(n_i) \underset{\textsc{License}}{\longrightarrow} \sum_{j=1}^N r_{j\to i} f_\mathrm{sc}^{(i)}(n_j),
\end{align}
where there are $N$ players, all engaging with one another using licensing only. Here, $r_{i\to j}$ denotes the proportion of $i$'s compute-time licensed to $j$, which are normalised (assuming no wasted clock-cycles) for all users such that,
\begin{align}
\sum_{j=1}^N r_{i\to j} = 1\,\,\forall\,i.	
\end{align}
Note that when \mbox{$r_{i\to j}=\delta_{i,j}$} (i.e \mbox{$r=I_N$} is the \mbox{$N\times N$} identity matrix, and there is no inter-player licensing) the \textsc{License} strategy simply reduces back to the \textsc{Isolation} strategy.

@@

Elements associated with \textsc{Unify} strategies will undergo the quantum utility payoff enhancement,
\begin{align}\label{eq:util_payoff_enh_1}
(f^{(X)}_\mathrm{sc}(n_X),f^{(Y)}_\mathrm{sc}(n_Y)) \underset{\textsc{Unify}}{\longrightarrow} (c_X,c_Y),
\end{align}
where,
\begin{align}\label{eq:util_payoff_enh_2}
c_X &= \frac{n_X}{n_X+n_Y}\cdot f^{(X)}_\mathrm{sc}(n_X+n_Y),\nonumber\\
c_Y &= \frac{n_Y}{n_X+n_Y}\cdot f^{(Y)}_\mathrm{sc}(n_X+n_Y),
\end{align}
from the definition for time-shared compute power given in Def.~\ref{def:time_share_comp_power}.

As before, we can logically generalise the payoff enhancement of a generalised \textsc{Unify} strategy to the multi-player scenario as,
\begin{align}
	f_\mathrm{sc}^{(i)}(n_i) \underset{\textsc{Unify}}{\longrightarrow} \frac{n_i}{\sum_{j=1}^N n_j} \cdot f_\mathrm{sc}^{(i)}\left(\sum_{j=1}^N n_j\right).
\end{align}

Thus, it is evident that the enhancement in \textsc{Unify} strategies is highly dependent on:
\begin{itemize}
	\item The total number of qubits held between the players,
		\begin{align}
			n_\mathrm{total} = \sum_j n_j.
		\end{align}
	\item The proportion of the qubits held by each player,
		\begin{align}
			r_i = \frac{n_i}{\sum_{j=1}^N n_j}.
		\end{align}
	\item The respective algorithms to which the computational resources are being applied by each player, which influence the player-specific subjective scaling functions, $f^{(i)}_\mathrm{sc}$, independently.
\end{itemize}

@@

What happens with collaboration between players who have different systems? If this assumption is relaxed, what comes of the resulting utility payoff matrix and choice of strategies? The analysis for this assumption had already been presented in the previous discussion. The effective impact of taxation is that it will limit the amount of computational power delivered to market and capture governmental revenue from the sale of compute-time. This is analogous to what happens if $X$ was to build a less powerful system than $Y$. Through collaboration, they could still attain point $Q'$, shown in Fig.~\ref{fig:game_theory_4}, but this time the asymmetry arises as a design consequence rather than a regulatory imposition. Despite $X$ now not contributing as much to the overall computational power of the quantum cloud, choosing the maximising strategy of offering all available computing power to the cloud does maximise individual utility. Furthermore, the subjective nature of the utility derivation means there doesn't have to be a change from the described payoff matrix shown in Fig.~\ref{fig:game_theory_3}. Despite having a less powerful system, the proportional allocation of financial rewards relative to the contribution, means both $X$ and $Y$ could maintain a symmetric utility payoff outcome, like in Fig.~\ref{fig:game_theory_3}. In summary, the 2-player game shows that cooperation will result in the better outcome for both players.

\latinquote{Ceretis paribus.}

### 2019-02-07 Sections/the_quantum_space_race.tex
Added authors to sections
Quotes, editing
SH corrections

@@

Involving ground-based quantum network nodes is not an inherent problem as 80\% of the atmosphere by mass is located within the first 12km in altitude, yielding low effective thickness of the atmosphere ($\sim10$km)\index{Effective atmospheric thickness}. This effective thickness of course changes with the azimuth of the satellite in the sky. Near the horizon the line-of-sight\index{Line-of-sight} between satellite and ground-station is far greater than when the satellite is directly overhead. The light trajectory beyond this effective thickness regime effectively passes through vacuum with attenuation\index{Photon loss} limited only by diffraction\index{Diffraction}. Thus, when directly overhead, a satellite-to-ground\index{Satellites!Satellite-to-ground communication} channel is more favourable in terms of signal attenuation than long-distance ground-to-ground\index{Ground-to-ground communication} channels.

>>>>

\sectionby{Tim Byrnes}

\famousquote{The Earth is the cradle of humanity, but mankind cannot stay in the cradle forever.}{Konstantin Tsiolkovsky}
\newline

@@

Involving ground-based quantum network nodes is not prohibitive as 80\% of the atmosphere by mass is located within the first 12km in altitude, yielding low effective thickness of the atmosphere ($\sim10$km)\index{Effective atmospheric thickness}. Despite this apparently low exposure to attenuation, the loss is nonetheless substantial, on the order of $\sim$40dB (depending on optical frequency) when the satellite is directly overhead. This significantly reduces effective count rates, a problem that does not represent any kind of engineering limitation, but rather an inherent physical one that cannot foreseeably be mitigated (it seems implausible that much can possibly be done to circumvent the atmosphere in satellite-to-ground communication!). This effective thickness of course changes with the azimuth of the satellite in the sky. Near the horizon the line-of-sight\index{Line-of-sight} between satellite and ground-station is far greater than when the satellite is directly overhead. The light trajectory beyond this effective thickness regime effectively passes through vacuum with attenuation\index{Photon loss} limited only by diffraction\index{Diffraction}. Thus, when directly overhead, a satellite-to-ground\index{Satellites!Satellite-to-ground communication} channel is more favourable in terms of signal attenuation than long-distance ground-to-ground\index{Ground-to-ground communication} channels over free-space\index{Free-space}.
